Indoor Dining Shutdowns Negatively Affect Toronto Restaurants’ Performance and Business Practices

### 2021-02-27 outputs/paper/paper.Rmd
Edited more spelling and grammar issues

This study was conducted for three months, from October 1st to December 31st, 2020 where we studied the comparison between a control group and an intervention group. The intervention group had to partake in the study by their business as a take-out and delivery only restaurant, while the control group remained fully operational with indoor dining services available. This was done intentionally to maintain the integrity of the study by studying the differences between restaurants that closed their indoor dining services and the ones that did not. Survey data was collected via Google Forms to measure the effects of the intervention on restaurant’s employee numbers and revenue, and if changes were made in menu prices and operation days and hours during the intervention period. Survey data were processed and analyzed in R [@citeR], and a complete list of packages and their usages can be found in the Appendix. Based on survey responses, it was revealed that most restaurants in the intervention group who shut down their indoor dining services experienced a decrease in their total sales revenue compared to the previous quarter, with many reporting an overall decline of 50% or more. During the study period, restaurants in the intervention group made several changes to their restaurant practices to account for these changes such as reducing the number of employees, adjusting their menu prices, and reducing their hours and days of operation. Meanwhile, the restaurants in the control group who continued operating at full capacity remained relatively stable in terms of their revenue, and did not make any significant changes to their employee counts, menu prices, or operational periods. This suggests that shutdowns of indoor dining practices due to COVID-19 have large implications on restaurant performance, where the revenue, hours and days of operation, and menu prices all get impacted to adjust to deviations from the restaurant’s original business model.
```{r}

@@

Restaurants in the intervention condition received $1200.00 CAD for participating. This amount was determined based on funding allocations for this study. We acknowledged that participating in the intervention may result in some financial losses and put some businesses at risk, and wanted to provide some form of compensation to help alleviate those concerns as well as incentivize restaurants to participate. However, we also understand that this amount of money carries different weight for different types of restaurants. We provided resources and encouraged restaurants to look into external relief funding to support their businesses throughout and after the intervention period.

@@

In order for us to make causal inferences about the effects of our intervention, steps had to be taken to ensure we had a valid comparison group to compare the effects of the intervention to. One characteristic of a valid comparison group is that it has roughly the same characteristics as the intervention group [@gertler2016impact]. The Toronto restaurant scene is extremely diverse in terms of the size of the restaurants, the type of food served, and the target populations. While we wanted to cover a wide range of restaurant types, it was difficult to control for all these factors given the number of restaurants in Toronto. We relied on the process of randomized sampling and random assignment to ensure, at the very least, there were no systematic differences between groups since every restaurant had an equal chance of being assigned to either condition. Similarly, using stratification to control for location helped to ensure the groups were similar based on the number of restaurants picked from each borough. However, while the sampling method was selected to reduce differences between the groups, there may be significant differences between restaurants who choose to participate in the intervention versus those who do not which need to be addressed when discussing the effects of the intervention.

@@

Survey questions pertaining to the characteristics of the restaurants, such as location, size, type, and age of restaurants Figure \@ref(fig:samplechar) and Figure \@ref(fig:Q10) show us that:

* The largest concentration (nearly 60%) of restaurants in the Greater Toronto Area are centralized in downtown Toronto (28.26%), North York (15.78%), and Scarborough (13.74%)
* Most restaurants are of medium or large scale^[The survey asked respondents to consider their restaurant scale in terms of space and seating capacity, however the perceptions of scale may be subjective]
* The majority of restaurants who responded are relatively new (Have been operating for 3 years or less)
* Fast-food/quick service restaurants were the most common types, followed by casual dining, and then fine dining
* During the intervention period, restaurants received the most orders via UberEats, followed by DoorDash, and then other food courier services
* A small portion of restaurants utilized their own in-house delivery services, most of which were in the control group
* The distribution of these characteristics were roughly equal among restauraunts in both groups, meaning their sample characteristics are similar

@@

Table \@ref(tab:Q5Q6tTest) displays changes in the numbers of employees working at restaurants in the sample prior to the intervention period (Oct 1) and following the intervention period (Dec 31). This shows us that initially, both the control and treatment group had a very similar number of employees on average. Following the intervention period, however, the treatment group appears to have 10.26 fewer employees on average. Two-tailed Welch's t-tests comparing the two groups' mean employees shows that the difference prior to the intervention period is not a significant finding ( *t*(330) = -8.56, *p* > .05)., but the difference in employee numbers between the two groups post-intervention is significant( *t*(330) = 19.46, *p* < .05).

@@

* Most of the restaurants in the control group did not make any alterations to their menu prices, while most restaurants in the intervention group made alterations by raising or lowering prices, more of which chose to reduce their menu prices
* Most of the restaurants in the control group did not adjust their hours or days of operation, while most of the ones in the intervention group reduced their operating schedules
* For restaurants who reduced their hours of operation, most restaurants reduced their hours by 15%-30%, followed by 30%-50%
* Most restaurants who reduced their days of operation did so by 1-2 days

Pearson’s Chi-squared (*X*^2^) tests were conducted to see if there is a statistically significant difference between responses regarding restaurant adjustments in the intervention and control conditions. The test results show that the p-value of each of those adjustments is less than 0.001, meaning there is a significant relationship between condition (whether not not indoor dining services were shut down or not) and adjustments made to menu prices, hours of operation, and days of operations. That is, restaurants in the intervention condition were more likely to change their menu prices and reduce their operational hours and days.

@@

Proponents of the ghost kitchen model believe take-out and delivery focused businesses are here to stay as demands for food delivery are on the rise. [@Sundar2019]. Some may argue that the growth in popularity of delivery services, even before the pandemic, reflect changes in the population’s dining habits. The fast-paced society we live in may place a lot more value on the ease and convenience of food delivery than on dining inside a restaurant, which raises the question as to whether ghost kitchens are able to replace indoor dining services entirely. With that being said, while Torontonians enjoy the convenience of take-out and delivery, there is still value in keeping indoor dining experiences. The decreases in revenue observed in the intervention group suggest a significant portion of the restaurants previous source of revenue came from indoor dining. These restaurants had to adjust aspects of their business practices to cater to the intervention. By shutting down indoor dining and only allowing for deliveries and takeout orders, the restaurants had to make significant changes to their employee counts, menu price, and opening times. Despite these alterations, restaurants in the treatment group still had a decrease in their revenue compared to the previous quarter, while revenue in the control group remained stable. Based on this result, we can make an assumption that the dining habits of Torontontonians did not change even though the restaurants switched to a business model which highlighted the convenience of take-out and delivery. In fact, Torontonian’s require a mix of delivery services, takeout orders, and dining-in experiences at restaurants. Due to this reason, restaurants in the treatment group noticed a decline in their revenue during the time of the intervention. People want to dine-in at restaurants sometimes, so if they desire to dine-in and restaurants do not have that option then they will go to a restaurant which can offer those services to them.

@@

We would like to thank the Ontario Government for generously providing Petit Poll with $240,000 CAD in funding to carry out this study. The funding was used to provide compensation to restaurants for participating in the intervention. A sample size of 200 for the intervention condition was selected based on this amount, in which each restaurant who participated in the intervention received $1200. A total of 150 restaurants participated in the intervention and received payment, making the total amount of funding spent during the study $180,000 CAD

>>>>

This study was conducted for three months, from October 1st to December 31st, 2020 where we studied the comparison between a control group and an intervention group. The intervention group had to partake in the study by their business as a take-out and delivery only restaurant, while the control group remained fully operational with indoor dining services available. This was done intentionally to maintain the integrity of the study by studying the differences between restaurants that closed their indoor dining services and the ones that did not. Survey data was collected via Google Forms to measure the effects of the intervention on restaurant’s employee numbers and revenue, and if changes were made in menu prices and operation days and hours during the intervention period. Responses were processed and analyzed in R [@citeR], and a complete list of packages and their usages can be found in the Appendix. Based on survey responses, it was revealed that most restaurants in the intervention group who shut down their indoor dining services experienced a decrease in their total sales revenue compared to the previous quarter, with many reporting an overall decline of 50% or more. During the study period, restaurants in the intervention group made several changes to their restaurant practices to account for these changes such as reducing the number of employees, adjusting their menu prices, and reducing their hours and days of operation. Meanwhile, the restaurants in the control group who continued operating at full capacity remained relatively stable in terms of their revenue, and did not make any significant changes to their employee counts, menu prices, or operational periods. This suggests that shutdowns of indoor dining practices due to COVID-19 have large implications on restaurant performance, where the revenue, hours and days of operation, and menu prices all get impacted to adjust to deviations from the restaurant’s original business model.
```{r}

@@

Restaurants in the intervention condition received $1,200 CAD for participating. This amount was determined based on funding allocations for this study. We acknowledged that participating in the intervention may result in some financial losses and put some businesses at risk, and wanted to provide some form of compensation to help alleviate those concerns as well as incentivize restaurants to participate. However, we also understand that this amount of money carries different weight for different types of restaurants. We provided resources and encouraged restaurants to look into external relief funding to support their businesses throughout and after the intervention period.

@@

In order for us to make causal inferences about the effects of our intervention, steps had to be taken to ensure we had a valid comparison group to measure the effects of the intervention against. One characteristic of a valid comparison group is that it has roughly the same characteristics as the intervention group [@gertler2016impact]. The Toronto restaurant scene is extremely diverse in terms of the size of the restaurants, the type of food served, and the target populations. While we wanted to cover a wide range of restaurant types, it was difficult to control for all these factors given the number of restaurants in Toronto. We relied on the process of randomized sampling and random assignment to ensure, at the very least, there were no systematic differences between groups since every restaurant had an equal chance of being assigned to either condition. Similarly, using stratification to control for location helped to ensure the groups were similar based on the number of restaurants picked from each borough. However, while the sampling method was selected to reduce differences between the groups, there may be significant differences between restaurants who choose to participate in the intervention versus those who do not which need to be addressed when discussing the effects of the intervention.

@@

Survey questions pertaining to the characteristics of the restaurants, such as location, size, type, and age of restaurants (Figure \@ref(fig:samplechar) and Figure \@ref(fig:Q10)) show us that:

* The largest concentration (nearly 60%) of restaurants in the Greater Toronto Area are centralized in downtown Toronto (28.26%), North York (15.78%), and Scarborough (13.74%);
* Most restaurants are of medium or large scale^[The survey asked respondents to consider their restaurant scale in terms of space and seating capacity, however the perceptions of scale may be subjective.];
* The majority of restaurants who responded are relatively new (Have been operating for 3 years or less);
* Fast-food/quick service restaurants were the most common types, followed by casual dining, and then fine dining;
* During the intervention period, restaurants received the most orders via UberEats, followed by DoorDash, and then other food courier services;
* A small portion of restaurants utilized their own in-house delivery services, most of which were in the control group; and
* The distribution of these characteristics were roughly equal among restauraunts in both groups, meaning their sample characteristics are similar.

@@

Table \@ref(tab:Q5Q6tTest) displays changes in the numbers of employees working at restaurants in the sample prior to the intervention period (Oct 1) and following the intervention period (Dec 31). This shows us that initially, both the control and treatment group had a very similar number of employees on average. Following the intervention period, however, the treatment group appears to have 10.26 fewer employees on average. Two-tailed Welch's t-tests comparing the two groups' mean employees shows that the difference prior to the intervention period is not a significant finding \newline(*t*(330) = -8.56, *p* > .05), but the difference in employee numbers between the two groups post-intervention is significant(*t*(330) = 19.46, *p* < .05).

@@

* Most of the restaurants in the control group did not make any alterations to their menu prices, while most restaurants in the intervention group made alterations by raising or lowering prices, more of which chose to reduce their menu prices;
* Most of the restaurants in the control group did not adjust their hours or days of operation, while most of the ones in the intervention group reduced their operating schedules;
* For restaurants who reduced their hours of operation, most restaurants reduced their hours by 15%-30%, followed by 30%-50%; and
* Most restaurants who reduced their days of operation did so by 1-2 days.

Pearson’s Chi-squared (*X*^2^) tests were conducted to see if there is a statistically significant difference between responses regarding restaurant adjustments in the intervention and control conditions. The test results show that the p-value of each of those adjustments is less than 0.001, meaning there is a significant relationship between condition (whether or not indoor dining services were shut down or not) and adjustments made to menu prices, hours of operation, and days of operations. That is, restaurants in the intervention condition were more likely to change their menu prices and reduce their operational hours and days.

@@

Proponents of the ghost kitchen model believe take-out and delivery focused businesses are here to stay as demands for food delivery are on the rise. [@Sundar2019]. Some may argue that the growth in popularity of delivery services, even before the pandemic, reflect changes in the population’s dining habits. The fast-paced society we live in may place a lot more value on the ease and convenience of food delivery than on dining inside a restaurant, which raises the question as to whether ghost kitchens are able to replace indoor dining services entirely. With that being said, while Torontonians enjoy the convenience of take-out and delivery, there is still value in keeping indoor dining experiences. The decreases in revenue observed in the intervention group suggest a significant portion of the restaurants previous source of revenue came from indoor dining. These restaurants had to adjust aspects of their business practices to cater to the intervention. By shutting down indoor dining and only allowing for deliveries and takeout orders, the restaurants had to make significant changes to their employee counts, menu price, and opening times. Despite these alterations, restaurants in the treatment group still had a decrease in their revenue compared to the previous quarter, while revenue in the control group remained stable. Based on this result, we can make an assumption that the dining habits of Torontonians did not change even though the restaurants switched to a business model which highlighted the convenience of take-out and delivery. In fact, Torontonians require a mix of delivery services, takeout orders, and dining-in experiences at restaurants. Due to this reason, restaurants in the treatment group noticed a decline in their revenue during the time of the intervention. People want to dine-in at restaurants sometimes, so if they desire to dine-in and restaurants do not have that option then they will go to a restaurant which can offer those services to them.

@@

We would like to thank the Ontario Government for generously providing Petit Poll with $240,000 CAD in funding to carry out this study. The funding was used to provide compensation to restaurants for participating in the intervention. A sample size of 200 for the intervention condition was selected based on this amount, in which each restaurant who participated in the intervention received $1,200. A total of 150 restaurants participated in the intervention and received payment, making the total amount of funding spent during the study $180,000 CAD
